---
title: "Chapter 2"
author: "Dr Ivan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2.4.3 Grid approximation
```{r gridApproximation}
num_point <- 500
# define grid
grid <- seq(from = 0, to = 1, length.out = num_point)

# define prior
prior_unif <- rep(1, num_point)
prior_step <- ifelse(grid < 0.5, 0, 1)
prior_sla <- exp(-5 * abs(grid - 0.5))

# compute likelihood at each value in grid
likelihood <- dbinom(6, size=9, prob = grid)

# compute product of likelihood and prior
unstd_posterior <- likelihood * prior_gau

# standardise the posterior, so it sums to 1
posterior <- unstd_posterior / sum(unstd_posterior)

# Display the posterior
plot( grid, posterior, type = "b", xlab = "probability of water", ylab = "posterior probability")
```


## 2.4.4 Quadratic approximation
```{r quadraticApproximation, eval = FALSE}
library(rethinking)

globe_qa <- quap(
  alist(
    W ~ dbinom(W + L, p), # binomial likelihood
    p ~ dunif(0, 1)       # uniform prior
  ),
  data = list(W = 6, L = 3)
)

# display summary
precis(globe_qa)
```
```{r analyticalCalculation}
# analytical calculation
W <- 6
L <- 3
curve( dbeta(x, W+1, L+1), from = 0, to = 1)

# quadratic approximation
curve(dnorm(x, 0.67, 0.16), lty = 2, add = TRUE)
```


## 2.4.5 Markov vhain monte Carlo
```{r MCMC, eval = FALSE}
n_samples <- 1000
p <- rep(NA, n_samples)
p[1] <- 0.5
W <- 6
L <- 3
for (i in 2:n_samples) {
  p_new <- rnorm(1, p[i - 1], 0.1)
  if (p_new < 0) p_new <- abs(p_new)
  if (p_new > 1) p_new <- 2 - p_new
  q0 <- dbinom(W, W+L, p[i-1])
  q1 <- dbinom(W, W+L, p_new)
  p[i] <- ifelse(runif(1) < q1 / q0, p_new, p[i - 1])
}

# Compare with analytical
dens(p, xlim = c(0,1))
curve(dbeta(x, W+1, L+1), lty=2, add = TRUE)
```

# 2.6 Practice
## Easy
2E1) (2) Pr(rain|Monday)
2E2) (3) The probability that it is Monday, given that it is raining
2E3) (1) Pr(Monday|rain)
2E4) Discussions are weird and too school like

## Medium
2M1)
```{r 2M1}
W <- c(3, 3, 5)
L <- c(0, 1, 2)

num_point <- 50
test_case <- 2
# define grid
grid <- seq(from = 0, to = 1, length.out = num_point)

# define prior
prior_unif <- rep(1, num_point)
prior_step <- ifelse(grid < 0.5, 0, 1)

# compute likelihood at each value in grid
likelihood <- dbinom(W[test_case], 
                     size = W[test_case] + L[test_case], 
                     prob = grid)

# compute product of likelihood and prior
unstd_posterior <- likelihood * prior_unif

# standardise the posterior, so it sums to 1
posterior <- unstd_posterior / sum(unstd_posterior)

# Display the posterior
plot( grid, posterior, type = "b", xlab = "probability of water", ylab = "posterior probability")
```

2M2) Just change above unstd_posterior to use prior_step
2M3) 
```{r 2M3}
prob_land_give_earth <- 0.3
prob_earth <- 0.5
prob_land <- 0.65
prob_earth_give_land <- prob_land_give_earth * prob_earth / prob_land
prob_earth_give_land
```

2M4) Card 1 = BB: B then B or B then B
     Card 2 = BW: B then W or W then B
     Card 3 = WW: W then W or W then W
Getting B on side one gives 3 total paths. 2 of them have B as next option, so prob = 2/3

2M5) Card 1 = BB: B then B or B then B
     Card 2 = BW: B then W or W then B
     Card 3 = WW: W then W or W then W
     Card 4 = BB: B then B or B then B
Getting B on side one gives 5 total paths. 4 have B on side 2, so prob = 4/5

2M6) Card 1 = BB: B then B or B then B 1 Way to pull
     Card 2 = BW: B then W or W then B 2 Ways to pull
     Card 3 = WW: W then W or W then W 3 Ways to pull
Getting B on side one gives equivalent of 4 total paths (2 from BB, 1 from BW that has 2 pulls) with 2 have B side 2, so prob = 2/4 = 0.5

2M7) Card 1 = BB: B then B or B then B
     Card 2 = BW: B then W or W then B
     Card 3 = WW: W then W or W then W
Ways to pull 2 cards: 1,2; 1,3; 2,1; 2,3; 3,1; 3,2
1,2: B(B)B(W), b(b)w(b), B(B)B(W), b(b)w(b)
1,3: b(b)w(w), b(b)w(w), b(b)w(w), b(b)w(w)
2,1: B(W)B(B), B(W)B(B), W(B)B(B), W(B)B(B)
2,3: b(w)w(w), b(w)w(w), W(B)W(W), W(B)W(W)
3,1: W(W)B(B), W(W)B(B), W(W)B(B), W(W)B(B)
3,2: W(W)B(W), W(W)W(B), W(W)B(W), W(W)B(W)
First card has black 12 options. Second card has white cuts to 8. Total choices with BB of those 8 is 6. prob = 6/8 = 0.75
## Hard
2H1) prob_twin_a = 0.1 prob_sing_a = 0.9
     prob_twin_b = 0.2 prob_sing_a = 0.8
     prob_a = 0.5 prob_b = 0.5
     2 births:
     a_twin_twin = 0.1 * 0.1 = 0.01
     a_twin_sing = 0.1 * 0.9 = 0.09
     a_sing_twin = 0.9 * 0.1 = 0.09
     a_sing_sing = 0.9 * 0.9 = 0.81
     b_twin_twin = 0.2 * 0.2 = 0.04
     b_twin_sing = 0.2 * 0.8 = 0.16
     b_sing_twin = 0.8 * 0.2 = 0.16
     b_sing_sing = 0.8 * 0.8 = 0.64
     x_twin_twin = a_twin_twin * prob_a + b_twin_twin * prob_b = 0.025
     Given saw twin first birth:
     p_twin2_twin1 = x_twin_twin / sum(x_twin_x) = 1/6 (0.025/0.15)
     
2H2) pr_a_twin = pr_twin_a * pr_a / pr_twin = 0.1 * 0.5 / 0.15 = 1/3

2H3) pr_a_twinsing = a_twin_sing * pr_a / pr_twinsing
     pr_twinsing = a_twin_sing * pr_a + b_twin_sing * pr_b = 0.125
     pr_a_twin_sing = 0.36
     
2H4) Ignoring birth data:
     pr_a_testA = pr_testA_a * pr_a / pr_testA
     pr_testA = pr_testA_a * pr_a + pr_testA_b * pr_b
              = 0.8 * 0.5 + 0.35 * 0.5 = 0.575
     pr_a_testA = 0.8 * 0.5 / 0.575 = 0.6956522
     With birth data
     pa_a_testA = 0.8 * 0.36 / (0.8 * 0.36 + 0.35 * 0.64)
                = 0.5625
     
